---
title: "Assignment"
author: "Chentong Sun"
date: "2022-08-13"
output: html_document
---

## Setup

```{r}
library(readr)
library(ggplot2)
library(lubridate)
library(dplyr)
```

This Rmd file is for the Assignment in Data Science as a Field class. In this assignment, the NYPD Shooting Incident Data were analyzed. A total of 6 plots were presented, and one statistics model was created.

```{r}
dataset=read_csv("NYPD_Shooting_Incident_Data__Historic_.csv")
dataset
```

## Visualization 1: Distribution of Victims' Age in all the Cases

The result suggests that most victims were between 18 and 44.

```{r}
dataset1=subset(dataset,dataset$VIC_AGE_GROUP!="UNKNOWN")
dataset1=dataset1[,12]
dataset1=table(dataset1)
dataset1=data.frame(dataset1)
p1=ggplot(data=dataset1, aes(x=VIC_AGE_GROUP, y=Freq))+
  geom_bar(stat="identity")+ggtitle("Amount of Shooting Cases and Victims' Age Group")+
  labs(y= "Amount of Shooting Cases", x = "Victims' Age Group")
p1
```

## Visualization 2: Distribution of Victims' Sexuality in all the Cases

The result suggests that most victims were male.

```{r}
dataset2=subset(dataset,dataset$VIC_SEX!="U")
dataset2=dataset2[,13]
dataset2=table(dataset2)
dataset2=data.frame(dataset2)
p2=ggplot(data=dataset2, aes(x=VIC_SEX, y=Freq))+
  geom_bar(stat="identity")+ggtitle("Amount of Shooting Cases and Victims' sexuality")+
  labs(y= "Amount of Shooting Cases", x = "Victims' sexuality")
p2
```

## Modeling: Distribution of Cases Occurring Time in all the Cases

In this plot, amount of shooting cases in every hour of a day were presented. The result suggests that most victims were murdered in the night.

```{r}
dataset3=dataset[,3]
time=list()
for(i in dataset3){
  i=as.POSIXct(i, format = "%H:%M:%S")
  i=round.POSIXt(i,"hours")
  i=strftime(i, format="%H")
  time=c(time,i)
}
dataset3=data.frame(matrix(unlist(time), nrow=length(time), byrow=TRUE))
names(dataset3)[1]="Time"
dataset3=table(dataset3)
dataset3=data.frame(dataset3)

p3=ggplot(data=dataset3, aes(x=Time, y=Freq, group=1))+
  geom_line()+ggtitle("Amount of Shooting Cases and Time")+
  labs(y= "Amount of Shooting Cases", x = "Case Occur Time (h)")
p3
```

In this plot, all 24 hours were ranked based on the number of cases occurred, and the rank were plotted along with the number of cases.

```{r}
dataset3=dataset3[order(dataset3$Freq),]
dataset3=cbind(Rank= as.numeric(1:nrow(dataset3)), dataset3)

p4=ggplot(data=dataset3, )+
  geom_line(aes(x=Rank, y=Freq, group=1))+
  ggtitle("Amount of Shooting Cases and Time")+
  labs(y= "Amount of Shooting Cases", x = "Rank")
p4
```

The rank was modeled along with number of cases, and the predicted cases and actual cases were presented in the following plot. the red line is the predicted cases, and the blue line is the actual cases. Two lines looks pretty close, and the adjusted R-squared value is 0.9776.

```{r}
mod=lm(dataset3$Freq~dataset3$Rank, data=dataset3)
summary(mod)
dataset3=dataset3 %>% mutate(pred=predict(mod))

p5=ggplot(data=dataset3, )+
  geom_line(aes(x=Rank, y=Freq, group=1), color="blue")+
  geom_line(aes(x=Rank, y=pred, group=2), color="red")+
  ggtitle("The Predicted and Actual Amount of Shooting Cases versus Time")+
  labs(y= "Amount of Shooting Cases", x = "Rank")
p5
```

In this plot,rank of time was substituted by the time from X-axis. the red line is the predicted cases, and the blue line is the actual cases. Two lines were still closed. The model is reasonable.

```{r}
p6=ggplot(data=dataset3, )+
  geom_line(aes(x=Time, y=Freq, group=1), color="blue")+
  geom_line(aes(x=Time, y=pred, group=1), color="red")+
  ggtitle("The Predicted and Actual Amount of Shooting Cases versus Time")+
  labs(y= "Amount of Shooting Cases", x = "Case Occur Time (h)")
p6

```

```{r}
sessionInfo()
```

## **Identifying sources of bias**

In this assignment, I make analysis on number of shooting cases versus victims' gender, their age and the occurring time of cases. Although the information provided in the database are credible and accurate, but I think I still need some extra data to improve my proposition.

The gender and age composition of all the population in NY is important because a bigger population in a specific gender or age group could increase the chance for case occurring in this group. My analysis shows that people between 18 and 44 have a bigger risk of shooting case, but they are the main population group in the whole population, and they are more commonly exposed to the environment which shooting cases occurring. Therefore, their risk is actually smaller than the prediction.

Similarly, in my model, I predicted that the number of cases and rank of time are in a linear relationship. However, in the midnight, a much fewer population are exposed to the risk of shooting. Therefore, the actual risk in the night is much bigger than that in my model.

Therefore, a objective bias caused by the lack of information could influence the result.
